\documentclass[../mathNotesPreamble]{subfiles}

\begin{document}
%  \relscale{1.4} %TODO
  \section{6.1: Simple Interest}
  \begin{defn*}
    The \textbf{simple interest} $I$ is given by
      \[I=Prt\]
    where
      \begin{align*}
        I&=\textnormal{ interest (in dollars) }\\
        P&=\textnormal{ principal (in dollars) }\\
        r&=\textnormal{ annual interest rate (as a decimal)}\\
        t&=\textnormal{ time (in years)}
      \end{align*}
    \textbf{Note: } The time measurements of $r$ and $t$ must agree\\
    
    From this, the \textbf{future value} of simple interest is
      \[S=P+I=P+Prt=P(1+rt)\]
  \end{defn*}
  \pagebreak
  
  \begin{ex*}\mbox{}
    \begin{extasks}[after-item-skip=\stretch{1}](1)
      \task If $\$8,000$ is invested for 2 years at an annual rate of $9\%$, how much interest will be received at the end of the $2$--year period? What will the future value be?
      \task If $\$4,000$ is borrowed for 39 weeks at an annual interest rate of $15\%$, how much interest is due at the end of the $39$ weeks? 
      \task An investor wants to have $\$20,000$ in $9$ months. If the best available simple interest rate is $6.05\%$ per year, how much must be invested now?
    \end{extasks}
  \end{ex*}
  \vspace*{\stretch{1}}
  \pagebreak

  \begin{defn*}
    The \textbf{return on investment} (ROI) is the ratio between the gain and cost of an investment:
      \[ROI=\frac{\textnormal{Gains on investment}}{\textnormal{Cost of investment}}\]
    The \textbf{earned (effective) interest rate} is the equivalent interest rate of the investment when all the fees and dividends are included.
  \end{defn*}
  \begin{ex*}
    Mary Spaulding bought Wind-Gen Electric stock for $\$6,125.00$. After $6$ months, the value of her shares had risen by $\$138.00$ and dividends totaling $\$144.14$ had been paid.
  \end{ex*}
  \begin{extasks}[after-item-skip=\stretch{1}](1)
    \task Find the return on investment on this investment.
    \task Find the simple interest rate she earned on this investment if she sold the stock at the end of the $6$ months.
  \end{extasks}
  \vspace*{\stretch{1}}
  \pagebreak
  
  \begin{ex*}\mbox{}
    \begin{extasks}[after-item-skip=\stretch{1}](1)
      \task To buy a Treasury bill (T-bill) that matures to $\$10,000$ in $6$ months, you must pay $\$9,750$. What annual simple interest rate does this earn?
      \task If the bank charges a fee of $\$40$ to buy a T-bill, what is the actual interest rate you earn?
    \end{extasks}
  \end{ex*}
  \vspace*{\stretch{1}}
  \pagebreak
  

  \pagebreak
\end{document}
